Examining the Impact of School Closures on COVID-19 Infections in Europe and their Effects on Different Age Cohorts

### 2024-02-20 manuscript_oup/manuscript.Rmd
oup

---
title: "Examining the Impact of School Closures on COVID-19 Infections in Europe and their Effects on Different Age Cohorts"
short_title: "Impact of School Closures on COVID-19 by Age Cohorts"
short_author: "Author 1 et. al."
authors: 
- name: "Author 1"
  affiliation: "Address"
  city: "City"
  country: "Country"
  email: "email1\\@example.com"
- name: "Author 2"
  affiliation: "Address2"
  city: "City2"
  country: "Country2"
  email: "email2\\@example.com"
# short_author: "Dupré & Morgenroth"
# authors:
# - name: "Damien Dupré"
#   affiliation: "Business School, Dublin City University"
#   city: "Dublin"
#   country: "Ireland"
#   email: "damien.dupre\\@dcu.ie"
#   footnote: 1
# - name: "Edgar Morgenroth"
#   affiliation: "Business School, Dublin City University"
#   city: "Dublin"
#   country: "Ireland"
#   email: "edgar.morgenroth\\@dcu.ie"
#   footnote: 2
# footnote:
#   - code: 1
#     text: "Corresponding Author"
#   - code: 2
#     text: "Equal contribution"
abstract: |
  This paper provides a thorough analysis of COVID-19 case trends in selected European countries, focusing on the influence of different age groups. It explores the relationship between population size and confirmed case numbers, suggesting that larger countries tend to have higher case figures. The study investigates how various waves of the pandemic affect different age demographics, with older age groups being most affected by the initial wave and younger age groups more impacted by subsequent waves.
  Using a Generalised Additive Model (GAM), the research assesses the impact of school closures on COVID-19 cases, revealing a decreasing non-linear effect across all countries examined. Following school closures, distinct trends emerge among age groups: ages 0 to 4 experience a decrease, ages 5 to 14 see an increase, and ages 15 to 24 initially surge then decline. Transfer Entropy calculations uncover imbalances in age group influences, with certain age groups predicting changes in others.
  Consistent patterns across countries are identified, such as younger cohorts influencing older cohorts in Austria, Germany, and the Netherlands, while Portugal and Spain show the opposite trend. These findings deepen our understanding of COVID-19 dynamics in Europe, providing valuable insights for targeted public health strategies and interventions.
keywords: COVID-19, School Closures, Causality, Generalised Additive Model, Transfer Entropy
bibliography: "bibliography.bib"
header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{threeparttable}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \usepackage{caption}

output:
  bookdown::pdf_book:
    base_format: rticles::rss_article
    keep_tex: false
  bookdown::word_document2:
    default
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("output"),
      output_file = c(glue::glue("{xfun::sans_ext(input)}.pdf"), glue::glue("{xfun::sans_ext(input)}.docx")),
      output_format = "all",
      envir = globalenv()
    )
  })
---

@@

@@

```{r te, fig.height=8, out.width="\\textwidth", fig.cap="(ref:te)"}
te_plot
```

@@

<div id="refs"></div>

>>>>

---
title: "Examining the Impact of School Closures on COVID-19 Infections in Europe and their Effects on Different Age Cohorts"
journal_title: "Journal of the Royal Statistical Society, Series C (Applied Statistics)"
short_title: "Impact of School Closures on COVID-19 by Age Cohorts"
short_author: "Dupré & Morgenroth"
authors_1_name: "Damien Dupré"
authors_1_ORCID: "0000-0001-8610-1045"
authors_2_name: "Edgar Morgenroth"
authors_2_ORCID: "0000-0002-9442-0561"
authors_affiliation: "Business School, Dublin City University"
authors_city: "Dublin"
authors_country: "Ireland"
corresponding_email: "damien.dupre\\@dcu.ie"
abstract: |
  This paper provides a thorough analysis of COVID-19 case trends in selected European countries, focusing on the influence of different age groups. It explores the relationship between population size and confirmed case numbers, suggesting that larger countries tend to have higher case figures. The study investigates how various waves of the pandemic affect different age demographics, with older age groups being most affected by the initial wave and younger age groups more impacted by subsequent waves.
  Using a Generalised Additive Model (GAM), the research assesses the impact of school closures on COVID-19 cases, revealing a decreasing non-linear effect across all countries examined. Following school closures, distinct trends emerge among age groups: ages 0 to 4 experience a decrease, ages 5 to 14 see an increase, and ages 15 to 24 initially surge then decline. Transfer Entropy calculations uncover imbalances in age group influences, with certain age groups predicting changes in others.
  Consistent patterns across countries are identified, such as younger cohorts influencing older cohorts in Austria, Germany, and the Netherlands, while Portugal and Spain show the opposite trend. These findings deepen our understanding of COVID-19 dynamics in Europe, providing valuable insights for targeted public health strategies and interventions.
keywords: COVID-19, School Closures, Causality, Generalised Additive Model, Transfer Entropy
bibliography: "bibliography.bib"
csl: elsevier-harvard.csl
header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{threeparttable}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \usepackage{caption}

output:
  bookdown::pdf_book:
    template: oup-authoring-template.tex
    keep_tex: false
---

@@

```{r te, fig.height=8, out.width="\\textwidth", fig.cap="(ref:te)", fig.pos = 'H'}
te_plot
```

@@

@@

<!-- <div id="refs"></div> -->
